\pkg{systemfit}: A Package for Estimating Systems of Simultaneous Equations in \proglang{R}

### 2008-09-12 paper/trunk/systemfit.Rnw
moved appendix section "Estimating systems of equations with sem" (systemfit_sem.Snw) into master file (systemfit.Rnw)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@960 591db2bd-e0a0-4d22-8c19-2192dbb7da31

\input{systemfit_sem.tex}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section[Estimating systems of equations with sem]
   {Estimating systems of equations with \code{sem}}
\label{sec:sem}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\SweaveOpts{prefix.string=trash/sem}
<<echo=FALSE>>=
options( prompt = "R> " )
options( width = 75 )
@

This section compares the commands to estimate a system of equations
by \code{sem} and \code{systemfit}.
This comparison uses Klein's °Model I° (see Section~\ref{sec:KleinsModel}).
Before starting the estimation,
we load the \pkg{sem} and \pkg{systemfit} package
as well as the required data set.
<<systemfit_sem>>=
library( "sem" )
library( "systemfit" )
data( "KleinI" )
@

First, we estimate the system by limited information maximum likelihood
(LIML) with \code{sem}:
<<systemfit_sem,keep.source = TRUE>>=
limlRam <- matrix(c(
   "consump  <-  corpProf",    "consump_corpProf",    NA,
   "consump  <-  corpProfLag", "consump_corpProfLag", NA,
   "consump  <-  wages",       "consump_wages",       NA,
   "invest   <-  corpProf",    "invest_corpProf",     NA,
   "invest   <-  corpProfLag", "invest_corpProfLag",  NA,
   "invest   <-  capitalLag",  "invest_capitalLag",   NA,
   "privWage <-  gnp",         "privWage_gnp",        NA,
   "privWage <-  gnpLag",      "privWage_gnpLag",     NA,
   "privWage <-  trend",       "privWage_trend",      NA,
   "consump  <-> consump",     "s11", NA,
   "privWage <-> privWage",    "s22", NA,
   "invest   <-> invest",      "s33", NA),
   ncol = 3, byrow = TRUE)
class(limlRam) <- "mod"
exogVar <- c("corpProf", "corpProfLag", "wages", "capitalLag", "trend",
   "gnp", "gnpLag")
endogVar <- c("consump", "invest", "privWage")
allVar <- c(exogVar, endogVar)

limlResult <- sem(ram = limlRam, S = cov(KleinI[ -1, allVar ]),
   N = (nrow(KleinI) - 1), fixed.x = exogVar)
print(limlResult)
@

Theoretically, the LIML results should be identical to OLS results.
Therefore, we re-estimate this model by OLS with \code{systemfit}.
<<systemfit_sem,keep.source=TRUE>>=
eqConsump <- consump ~ corpProf + corpProfLag + wages
eqInvest <- invest ~ corpProf + corpProfLag + capitalLag
eqPrivWage <- privWage ~ gnp + gnpLag + trend
system <- list(consump = eqConsump, invest = eqInvest,
   privWage = eqPrivWage)
olsResult <- systemfit(system, data = KleinI)
print(olsResult)
@
As expected, the results are identical.

Now, we estimate the system by full information maximum likelihood
(FIML) with \code{sem}:
<<systemfit_sem,keep.source = TRUE>>=
fimlRam <- rbind(limlRam,
   c("consump  <-> invest",   "s12", NA),
   c("consump  <-> privWage", "s13", NA),
   c("privWage <-> invest",   "s23", NA))
class(fimlRam) <- "mod"

fimlResult <- sem(ram = fimlRam, S = cov(KleinI[ -1, allVar ]),
   N = (nrow(KleinI) - 1), fixed.x = exogVar)
print(fimlResult)
@

Theoretically, results of an iterated SUR estimation should
converge to FIML results.
Therefore, we re-estimate this model by iterated SUR with \code{systemfit}.
<<systemfit_sem>>=
surResult <- systemfit( system, method = "SUR", data = KleinI,
   methodResidCov = "noDfCor", maxit = 500 )
print( surResult )
@
As expected, the results are rather similar.
